---
title: "Madies Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Madies Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## introducing Madies package

# This package will be able to do some basic cleaning, graphing, a conversion and a little bit of math (ANOVA and log). The idea behind this project is cleaning data for processing and further analyses and providing basic functions that will setup data for further analyses. 
 
## installation instructions
```{r setup}
devtools::install_github("mdenman00/Project_3")
library(MadiesPackageR)
```

## where to get the data 

+ could use google sheets and give people instructions
+ can stash data in R package

## using function one
``` {r}
# This will create an average for a column of your choice, if the average is less than the number you put it in you'll get a message back saying "too small"

MadiesPackageR::gutselli_function("data/clean_gutselli.csv", "SL", 20)
```

## function two
``` {r}
# this will log transform a column and create a new column of the log data

MadiesPackageR::log_transform()
```

#function three
``` {r}
# creates an ANOVA of data and provides different plots of ANOVA

MadiesPackageR:: anova_function()
```

#function four
``` {r}
# user can create a point, line or bar graph and set their parameters for each 

MadiesPackageR::point_line_bar()
```

#function five
```{r}
# this will omit NAs, user has to set what they want to the change the NAs to be (ex: blank space, -,...)

MadiesPackageR::clean_data()
```

#function six
```{r}
# converts centimeters to inches in a given column

MadiesPackageR::inches()
```

